---
title: "District IV Profile"
author: "David McClendon"
date: "9/15/2019"
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
---
<style type="text/css">
body { font-size: 16px; line-height:140%: font-family: 'Georgia', sans-serif; color:#333; }
.tocify { font-size: 12px; border:none!important; border-radius:0px!important; }
h2 { font-size:30px; }
#TOC {
  background: url("https://images.squarespace-cdn.com/content/5b4f6c5bcef3723dc9e26070/1539201004682-4EGL1XPJ9UTELJ407TNO/LISDLOGO_horizontal_NEW_dark.png?format=1500w&content-type=image%2Fpng");
  background-size: 80%;
  padding-bottom: 40px !important;
  background-repeat: no-repeat;
  background-position: left bottom;
}
.list-group-item.active, .list-group-item.active:focus, .list-group-item.active:hover {
background-color: #327DBF;
border-color: #327DBF;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
library(tidyverse)
library(leaflet)
library(DT)
library(jastyle)
library(sf)
library(sp)
library(htmlwidgets)

dfsf <- readRDS(here::here("getStarted.rds"))
df <- dfsf
df$geometry <- NULL
```

# District comparison
These tables compare District IV to other HISD Board districts on demographics and student achievement. All data comes from the Texas Education Agency (TEA).

### Race-ethnicity
Over half of students in District IV are Black/African American. This is the largest share of any board district.

```{r race, echo=F}
district_summary <- df %>% 
  group_by(NAME2) %>% 
  summarise(
    `Students` = sum(CPETALLC, na.rm = T),
    `% Asian-PI` = (sum(CPETASIC, na.rm = T) + sum(CPETPCIC, na.rm = T))/`Students`,
    `% Black` = sum(CPETBLAC, na.rm = T)/`Students`,
    `% Hispanic` = sum(CPETHISC, na.rm = T)/`Students`,
    `% NH White` = sum(CPETWHIC, na.rm = T)/`Students`,
    `% Other` = (sum(CPETINDC, na.rm = T) + sum(CPETTWOC, na.rm = T))/`Students`
  ) %>% 
  rename(`HISD District` = NAME2)

datatable(district_summary, options = list(pageLength = 20, dom = 't'), rownames=F) %>% 
  formatPercentage(c("% Asian-PI","% Black","% Hispanic","% NH White", "% Other"), 0)

```

### Income and language
District IV has lowest share of LEP students (14%). 3-in-4 students are economically disadvantaged (low-income and receiving free or reduced-price lunch). This is on par with the district at large. 

```{r ecodis, echo=F}
district_summary <- df %>% 
  group_by(NAME2) %>% 
  summarise(
    `Students` = sum(CPETALLC, na.rm = T),
    `% Economically Disadvantaged` = sum(CPETECOC, na.rm = T)/`Students`,
    `% LEP students` = sum(CPETLEPC, na.rm = T) /`Students`

  ) %>% 
  rename(`HISD District` = NAME2)

datatable(district_summary, options = list(pageLength = 20, dom = 't'), rownames=F) %>% 
  formatPercentage(c("% Economically Disadvantaged","% LEP students"), 0)
```

### School accountability rating (A-F)
```{r}
accountability <- df %>% 
  count(NAME2, C_RATING) %>% 
  filter(!is.na(C_RATING)) %>% 
  group_by(NAME2) %>% 
  mutate(
    pct = n/sum(n)
  ) %>% 
  dplyr::select(-n) %>% 
  spread(C_RATING, pct) %>% 
  rename(`HISD District` = NAME2)

datatable(accountability, options = list(pageLength = 20, dom = 't'), rownames=F) %>% 
  formatPercentage(c("A","B", "C", "D", "F", "Not Rated"), 0)

```
*Percentage of schools receiving A-F grade in 2018-2019*

# District IV map
```{r}
mapdf <- dfsf %>% filter(NAME2=="IV") %>% 
  mutate(
    GRDTYPE = ifelse(GRDTYPE=="E", "Elementary",
                     ifelse(GRDTYPE=="M", "Middle",
                            ifelse(GRDTYPE=="S", "High", "Middle/High"))),
    GRDTYPE = ifelse(is.na(GRDTYPE), "Early Childhood", GRDTYPE)
  )



schoolpal <- colorFactor(c("#A1B73B", "#1AB2F0", "#E5B235", "#D8382C", "purple"), n=5, domain = mapdf$GRDTYPE)
#Color based on elem, middle, high

leaflet(width = "100%") %>% 
  addProviderTiles("CartoDB.Positron", group="Greyscale") %>% 
  addCircleMarkers(data=mapdf, 
                   color = schoolpal(mapdf$GRDTYPE),
                   label = paste0(mapdf$School_Nam),
                   popup= ifelse(mapdf$School_Nam == "KANDY STRIPE ACADEMY",
                                 paste0("<strong>", mapdf$School_Nam, "</strong></br>",
                                        "Not open 2018-2019 school year."),
                                 paste0("<strong>", mapdf$School_Nam, "</strong></br>",
                                "Asian-Pacific Islander: ", mapdf$CPETASIP, "%</br>",
                                "Black/African American: ", mapdf$CPETBLAP, "%</br>",
                                "Hispanic: ", mapdf$CPETHISP, "%</br>",
                                "Non-Hispanic White: ", mapdf$CPETWHIP, "%</br>",
                                "Other race-ethnicity: ", mapdf$CPETTWOP + mapdf$CPETINDP, "%</br>",
                                "Economically disadvantaged: ", mapdf$CPETASIP, "%</br>",
                                "LEP students: ", mapdf$CPETASIP, "%</br>",
                                "2019 Accountability rating: ", mapdf$C_RATING
                                ))
                      ) %>%
  addLegend("bottomright", pal = schoolpal, 
            values = mapdf$GRDTYPE,
            title = "Grade level",
            opacity = 0.8)
```



# Elementary schools
## Race-ethnicity
```{r}
df4 <- df %>% 
  filter(NAME2=="IV") %>% 
  mutate(
    `School` = School_Nam,
    `Students` = CPETALLC,
    `% Asian-PI` = CPETASIP/100 + CPETPCIP/100,
    `% Black` = CPETBLAP/100,
    `% Hispanic` = CPETHISP/100,
    `% NH White` = CPETWHIP/100,
    `% Other` = CPETINDP/100 + CPETTWOP/100,
    `% Economically Disadvantaged` = CPETECOP/100,
    `% LEP` = CPETLEPP/100,
    `Overall` = C_RATING,
    `Student acheivement` = CD1G,
    `School progress` = CD2AG,
    `Closing the gaps` = CD3G
  )

race <- df4 %>% 
  filter(GRDTYPE == "E") %>% 
  dplyr::select(`School`, `Students`:`% Other`) %>% 
  arrange(`School`)

datatable(race, options = list(pageLength = 20, dom = 't'), rownames=F) %>% 
  formatPercentage(c("% Asian-PI","% Black","% Hispanic","% NH White", "% Other"), 0)

```


## Income and language
```{r}
inclang <- df4 %>% 
  filter(GRDTYPE == "E") %>% 
  dplyr::select(`School`, `% Economically Disadvantaged`:`% LEP`) %>% 
  arrange(`School`)

datatable(inclang, options = list(pageLength = 20, dom = 't'), rownames=F) %>% 
  formatPercentage(c("% Economically Disadvantaged","% LEP"), 0)
```


## School Accountability
```{r}
acct <- df4 %>% 
  filter(GRDTYPE == "E") %>% 
  dplyr::select(`School`, `Overall`:`Closing the gaps`) %>% 
  arrange(`School`)

datatable(acct, options = list(pageLength = 20, dom = 't'), rownames=F) 
```


# Middle schools
## Race-ethnicity
```{r}
race <- df4 %>% 
  filter(GRDTYPE == "M") %>% 
  dplyr::select(`School`, `Students`:`% Other`) %>% 
  arrange(`School`)

datatable(race, options = list(pageLength = 20, dom = 't'), rownames=F) %>% 
  formatPercentage(c("% Asian-PI","% Black","% Hispanic","% NH White", "% Other"), 0)

```


## Income and language
```{r}
inclang <- df4 %>% 
  filter(GRDTYPE == "M") %>% 
  dplyr::select(`School`, `% Economically Disadvantaged`:`% LEP`) %>% 
  arrange(`School`)

datatable(inclang, options = list(pageLength = 20, dom = 't'), rownames=F) %>% 
  formatPercentage(c("% Economically Disadvantaged","% LEP"), 0)
```


## School Accountability
```{r}
acct <- df4 %>% 
  filter(GRDTYPE == "M") %>% 
  dplyr::select(`School`, `Overall`:`Closing the gaps`) %>% 
  arrange(`School`)

datatable(acct, options = list(pageLength = 20, dom = 't'), rownames=F) 
```


# High schools
## Race-ethnicity
```{r}
race <- df4 %>% 
  filter(GRDTYPE %in% c("S", "B")) %>% 
  dplyr::select(`School`, `Students`:`% Other`) %>% 
  arrange(`School`)

datatable(race, options = list(pageLength = 20, dom = 't'), rownames=F) %>% 
  formatPercentage(c("% Asian-PI","% Black","% Hispanic","% NH White", "% Other"), 0)

```


## Income and language
```{r}
inclang <- df4 %>% 
  filter(GRDTYPE %in% c("S", "B")) %>% 
  dplyr::select(`School`, `% Economically Disadvantaged`:`% LEP`) %>% 
  arrange(`School`)

datatable(inclang, options = list(pageLength = 20, dom = 't'), rownames=F) %>% 
  formatPercentage(c("% Economically Disadvantaged","% LEP"), 0)
```


## School Accountability
```{r}
acct <- df4 %>% 
  filter(GRDTYPE %in% c("S", "B")) %>% 
  dplyr::select(`School`, `Overall`:`Closing the gaps`) %>% 
  arrange(`School`)

datatable(acct, options = list(pageLength = 20, dom = 't'), rownames=F) 
```

